R Notebook

### 2018-03-12 paper/writing/structure.tex
its getting better...

- More information available, more pluralism? 
 Pluralism and diversely sourced media are on the decline, whereas concentration of media ownership and consolidation are increasing.

More specifically, I use a data set of German online news articles about domestic politics dated from 01.06.2017 to 31.12.2017\footnote{German federal elections took place on 24th of September 2017.} from six German news content providers. I analyze how the topics are discussed differently in public and private news.

The research strategy is as follows:

\begin{enumerate}
	\item \textbf{Discovering Topics (Section \ref{ch_model})} To discover the latent topics in the corpus, the structural topic modeling (STM) developed by \citet{roberts_model_2016} is applied. The STM is an unsupervised machine learning approach that models topics as multinomial distributions of words and documents as multinomial distributions of topics, allowing to incorporate external variables that effect both, topical content and topical prevalence. I will include the business model (public or private) as a control for both the topical content and the topical prevalence. Additional, the month an article was published is included as a control for the topical prevalence. The result of the generative process of STM are the posterior distributions for topic prevalence and topical content, which can then be used to measure the similarity between private and public news providers in the further course of the paper: (1) $\beta$ is a $K$-by-$V$ matrix (where $K=$ number of topics and $V=$ vocabulary), where the entry $\beta_{kvc}$ can be interpreted as the probability of observing the $v$-th word in topic $k$ for the covariate level $c$. (2) $\theta$ is a $D$-by-$V$ matrix of the document-topic distributions, where the entry $\theta_{dk}$ can be interpreted as the proportion of words in document $d$ which arise from topic $k$, or rather as the probability that document $d$ deals about topic $k$.

	\item \textbf{Measuring Similarity of news content} Different approaches are used to examine similarities and/or differences of content between the news provider:
	\begin{enumerate}
		\item\label{item_1}\textbf{Differences in topic prevalence:} As I allow prevalence of topics to evolve over time (by month) and vary across business models, the posterior topic-probability $\theta_{dk}$ of document can be used to estimate the conditional expectation of topic prevalence for given document characteristics.

		\item\label{item_2} \textbf{Topic correlation} To get an idea of how topics are linked and framed differently across news wires, the correlation of the posterior probability of topics $\theta_{dk}$ is calculated.

		\item\label{item_3} \textbf{Differences in word-topic distributions} As each row of $\beta_{kvc}$ corresponds to the word-distribution of a respective topic for the covariate level $c$, it can be used to calculate the similarity of the word-topic distribution between the news provider using established similarity measures.

@@

Approaches \ref{item_1} and \ref{item_2} have been used in \citet{roberts_model_2016}, among others, to examine differences in the content of news providers. However, I extend the analysis by calculating the similarity of the word-topic distribution between the news provider, to identify which topics are discussed similar or differently (\ref{item_3}). In addition, this approach is applied to a new dataset covering recent political trends in Germany.

The remaining course of the paper is as follows: In Section \ref{ch_onlinenews} I briefly discuss the characteristics of the market for online news in Germany. The data used to conduct the model is described in Section \ref{ch_data}. Section \ref{ch_model} explains the generative process of the structural topic model as well as the selected parameters to run the model. The results are the evaluated in Section \ref{ch_empirical}.

@@

I conduct the estimation on a sample of 9,393 online news articles from six news provider about domestic politics\footnote{Bild.de, DIE WELT, FOCUS ONLINE, SPIEGEL ONLINE, ZEIT ONLINE, Tagesschau.de}, of which only Tagesschau.de belongs to the group of public provider. The reason for this is that the content structure of Tagesschau.de is most similar to that of the private providers. ZDF.de offers predominantly video content and DLF website mainly offers audio content in the form of interviews, which makes it hard to include it in the model. However, as only a part of the private suppliers are included to maintain the proportions between private and public market participants are maintained, thus minimizing possible bias. The articles are dated from 01.06.2017 to 31.12.2017.\footnote{German federal elections took place on 24th of September 2017.} I first extract all online articles using the the Webhose.io API.\footnote{For more information see https://docs.webhose.io/v1.0/docs/getting-started. The scraping code was written in Python and can be made available on request.} Then all articles from the section "domestic policy" are filtered by checking the URL structure.

>>>>

At the same time, the number of people who find out about current news on the Internet is also increasing. According to a study on the media coverage of the German-speaking population, nearly 40\% of respondents used the internet at least once a week in 2017 to inform themselves about the current news compared to 34\% in 2016 \citep{vuma_arbeitsgemeinschaft_verbrauchs-_und_medienanalyse_verbrauchs-_2017}.

This paper examines to what extent the content of different online news resembles each other. More specifically, I use a data set of German online news articles about domestic politics dated from 01.06.2017 to 31.12.2017\footnote{German federal elections took place on 24th of September 2017.} from seven German news content providers. The final step is to check whether the tonality of reporting is reflected in the polls. The empirical strategy follows a novel approach combining "the two Ts": Topic and Tone \citep{hansen_shocking_2016}. That is, first the topics covered in the articles are identified, to find out how these topics are discussed (tone). More precisely, the research strategy is as follows:

\begin{enumerate}
	\item \textbf{Discovering Topics (Section \ref{ch_model})} To discover the latent topics in the corpus, the structural topic modeling (STM) developed by \citet{roberts_model_2016} is applied. The STM is an unsupervised machine learning approach that models topics as multinomial distributions of words and documents as multinomial distributions of topics, allowing to incorporate external variables that effect both, topical content and topical prevalence. I estimate a model, where the newswire service is included as a control for both the topical content and the topical prevalence. The result of the generative process of STM are the posterior distributions for topic prevalence and topical content, which can then be used to measure the similarity between the news providers in the further course of the paper: (1) $\beta$ is a $K$-by-$V$ matrix (where $K=$ number of topics and $V=$ vocabulary) containing the probability distribution for topic $k$ over the words in the predefined vocabulary. (2) $\theta$ is a $D$-by-$V$ matrix of the document-topic distributions, where the entry $\theta_{dk}$ can be interpreted as the proportion of words in document $d$ which arise from topic $k$, or rather as the probability that document $d$ deals about topic $k$. As the prevalence of topics are allowed to vary across news wires, the posterior topic-probability $\theta_{dk}$ of an article (document) can be used to estimate the conditional expectation of topic prevalence for given article characteristics.

	\item \textbf{Measuring Tone} After assigning a topic to each article (document), a dictionary-based sentiment analysis is conducted to estimate how topics are discussed differently by different news wires.

@@

Approach \ref{item_1} has been used in \citet{roberts_model_2016}, among others, to examine differences in the content of news providers. However, I extend the analysis by comparing the sentiment score for a given topic between the news provider, to identify which topics are discussed similar or differently (\ref{item_2}). \citet{hansen_shocking_2016} applied a similar approach to a dataset of 142 FOMC decision statements to measure the effect of those statements on macroeconomic variables.

The remaining course of the paper is as follows: In Section \ref{ch_onlinenews} the characteristics of the market for online news in Germany are outlined. The data used to conduct the model is described in Section \ref{ch_data}. Section \ref{ch_model} explains the generative process of the structural topic model as well as the selected parameters to run the model. The results of the STM, as well as the sentiment analysis are the evaluated in Section \ref{ch_empirical}.

@@

I conduct the estimation on a sample of 14,937 online news articles from six news provider about domestic politics\footnote{Bild.de, DIE WELT, FOCUS ONLINE, SPIEGEL ONLINE, Stern.de, ZEIT ONLINE, Tagesschau.de}, of which only Tagesschau.de belongs to the group of public provider. The reason for this is that the content structure of Tagesschau.de is most similar to that of the private providers. ZDF.de offers predominantly video content and DLF website mainly offers audio content in the form of interviews, which makes it hard to include it in the model. The articles are dated from 01.06.2017 to 01.03.2018.\footnote{German federal elections took place on 24th of September 2017.} I first extract all online articles using the the Webhose.io API.\footnote{For more information see https://docs.webhose.io/v1.0/docs/getting-started. The scraping code was written in Python and can be made available on request.} Then all articles from the section "domestic policy" are filtered by checking the URL structure.
